---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# seamonas

<!-- badges: start -->
<!-- badges: end -->

The goal of seamonas is to provide tools to make plots of the North Sea

# 1. Intro

Contains:

- Data from a random generated survey in CRS 3035 and CRS 4326  
- A grid generated using the function create_grid in CRS 3035

# 2. Installation

You can install the development version of seamonas from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("MiriamLL/seamonas")
```

```{r example, eval=TRUE}
library('seamonas')
```

# 3. Data

```{r}
head(survey_4326)
```

## 3.1. Survey data

```{r, eval=FALSE}
survey_4326<-survey_4326
```

```{r}
ggplot2::ggplot()+
    ggplot2::geom_point(data=survey_4326,
                        ggplot2::aes(x=longitude, y= latitude, color=date),size = 1, shape = 16) 
```
Other data included:

```{r, eval=FALSE}
grid5x5_3035
grid_surveyed
density_survey
```

# 4. Survey grid

## 4.1. transform_survey

If you collected latitude and longitude in degrees, transform from 4326 to 3035

```{r}
survey_3035<-transform_survey(survey_data=survey_4326,
                 column_latitude='latitude',
                 column_longitude='longitude',
                 from_CRS=4326,
                 to_CRS=3035)
```


## 4.2. subset_grid

Add values to geometries of the grid, each square is a geometry or polygon
There will be warning that variables are assumed to be spatially constant throughout all geometries
Will print a plot and return the grids with data

```{r, echo=FALSE, eval=FALSE}
grid_surveyed<-subset_grid(survey_grid=grid5x5_3035,
                      survey_data=survey_3035,
                      grid_identifier='grid_id')
```


## 4.3. plot to check grid

```{r, echo=FALSE, eval=FALSE}
ggplot2::ggplot()+
  ggplot2::geom_sf(data = grid_surveyed, colour = "#42a921", fill= '#bde0fe',alpha=0.9)+
  ggplot2::geom_sf(data = survey_3035)+
  NULL
```




# 5. Densities grid

## 5.1. load data

```{r}
grid_surveyed<-grid_surveyed
density_survey<-density_survey
```

## 5.2. transform_survey

```{r}
densities_3035<-transform_survey(survey_data=density_survey, 
                 column_latitude='latitude',
                 column_longitude='longitude',
                 from_CRS=4326,
                 to_CRS=3035)
```

## 5.3. grid_densities

```{r}
density_grid<-subset_density(density_survey=densities_3035,
                             column_density='densities',
                             survey_grid=grid_surveyed,
                             grid_identifier='grid_id')
```


## 5.4. plot to check grid

```{r}
ggplot2::ggplot()+
  ggplot2::geom_sf(data = density_grid,mapping = ggplot2::aes(fill = mean_density), lwd = 0, colour = NA)
```


# 6. Classify densities

## 6.1. plot_density

```{r}
my_CRS<-3035
Europa<-sf::st_transform(GermanNorthSea::German_land, my_CRS)
EEZ<-sf::st_transform(GermanNorthSea::German_EEZ, my_CRS)
color_land='#f7bf54'
color_water='#3668b4'
xval<-c(3910000,4250000)
yval<-c(3380000,3680000)
```


```{r}
density_plot<-density_plot<-ggplot2::ggplot()+
    # maps
    ggplot2::geom_sf(data = EEZ, colour = 'black', fill = color_water)+
    ggplot2::geom_sf(data = Europa, colour = 'black', fill = color_land)+
    ggplot2::geom_sf(data = density_grid,mapping = ggplot2::aes(fill = mean_density), lwd = 0, colour = NA) +
    ggplot2::coord_sf(xlim = xval, ylim = yval)+

    NULL
density_plot
```


## 6.2. add_breaks

```{r}
plot_wbreaks<-add_breaks(density_plot=density_plot,
                     legendbreaks=c(0,0.5,1,1.5,2.0),
                     legendlimits=c(0,2),
                     legendlabels=c('     0','   0.5','   1.0','   1.5','> 2.0'))
plot_wbreaks
```



## 6.3. add_legend

```{r}
plot_wlegend<-add_legend(
  plot_wbreaks=plot_wbreaks,
  legtx=3905000,
  legty=3510000,
  legxy=c(0.11, 0.21),
  xval=c(3910000,4250000),
  yval=c(3380000,3680000))
plot_wlegend
```


## 6.4. add_theme

```{r}
plot_wtheme<-add_theme(plot_wlegend = plot_wlegend)
plot_wtheme
```







