---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# seamonas

<!-- badges: start -->
<!-- badges: end -->

The goal of seamonas is to provide tools to make plots of the North Sea

# Intro

Contains data:  
- Euring data with taxonomic grouping  
Last update: Mar-2024  
Base information from November 2023 (https://euring.org/data-and-codes/euring-codes)  
Except 6169 was changed to 6159.


```{r, eval=FALSE, echo=FALSE}
# Correct after 2024, according to SQL data base
library(seamonas)
Code_Euring<-Code_Euring
## Check which code needs to be corrected
Code_Euring %>% filter(Code=='6169')
## Check that the code is not currently on use
Code_Euring %>% filter(Code=='6159')
## Substitute
Code_Euring$Code<-gsub('6169', '6159',Code_Euring$Code)
## Check that was correctly saved
Code_Euring %>% filter(Code=='6159')
# Check that it was removed
Code_Euring %>% filter(Code=='6169')
## Save
save(Code_Euring, file = "Code_Euring.RData", version = 2)
```

```{r, eval=FALSE, echo=FALSE}
library(seamonas)
Euring_codes1<-Code_Euring
Euring_codes1$English_name<-gsub("/", " / ",Euring_codes1$English_name) 

library(tidyverse)

library(readr)
Euring_codes2a <- read_csv("data/SpeciesCodes2020_20231114_0.csv")
colnames(Euring_codes2a)<-c('Code','Scientific_name2','Updated_Date')
Euring_codes2b<-Euring_codes2a %>% select(Code,Scientific_name2)

library(readxl)
Euring_codes3a<-read_excel("data/LKP_EURING.xlsx")
Euring_codes3b<-Euring_codes3a %>% select(CODE,SCIENTIFIC_NAME,ENGLISH_NAME)
Euring_codes3c<-Euring_codes3b %>% rename(Code=CODE, Scientific_name3=SCIENTIFIC_NAME,English_name3=ENGLISH_NAME)

Euring_codes4<-merge(Euring_codes1,Euring_codes2b,by='Code',all=TRUE)
Euring_codes5<-merge(Euring_codes4,Euring_codes3c,by='Code',all=TRUE)
Euring_codes5<-Euring_codes5 %>%
  select(Code,Scientific_name,Scientific_name2,Scientific_name3,English_name,English_name3)

Euring_codes6 <- Euring_codes5 %>%
  drop_na(Scientific_name3)

Euring_codes6 <- Euring_codes6 %>%
  mutate(Same_names=if_else(English_name == English_name3, "equal", "different"))
Euring_codes7<-Euring_codes6 %>%
  filter(Same_names=='different')
```


- Data from a random generated survey in CRS 3035 and CRS 4326   
- Data from random generated densities  
- A grid generated using the function create_grid in CRS 3035 

Contains functions for spatial analyses:  
- transform_survey  
- subset_grid  
- grid_densities  

Contains function to plot:  
- plot_density  
- add_breaks  
- add_legend  
- add_theme  



3# Installation

You can install the development version of seamonas from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("MiriamLL/seamonas")
```

```{r example, eval=TRUE}
library('seamonas')
```

# Data

Data randomly generated to simulate surveys, density data and data on grids

## classification by euring


```{r TaxonomicGroup, echo=FALSE, eval=TRUE}
ImportTaxonomicGroups<-function(Species){
Divers<-c("Black-throated Diver","Great Northern Diver","Red-throated Diver","unidentified diver")

#Combine cormorants and gannets
Cormorant_Gannets<-c("Great Cormorant","Great Cormorant/European Shag","Northern Gannet")

Anatidae<-c("Common Eider","Common Scoter","Eurasian Teal","Long-tailed Duck","Red-breasted Merganser","Eurasian Wigeon","Northern Shoveler","Velvet Scoter","unidentified duck","Mallard","Goosander","Common Pochard")

#Please move Mute Swan and Cmmon Shelduck to separate classification
Geese_Shelducks<-c("Brent Goose","Greylag Goose","Barnacle Goose","Canada Goose","Mute Swan","Common Shelduck")

Skuas<-c("Arctic Skua","Great Skua","Long-tailed Skua","Pomarine skua")

Auks<-c("Atlantic Puffin","Common Guillemot","Common Guillemot/Razorbill", "Razorbill","unidentified auk")

Grebes<-c("Great Crested Grebe","unidentified grebe","Black-necked Grebe","Red-necked Grebe")

Waders<-c("Eurasian Oystercatcher","European Golden Plover","Red Knot","unidentified wader","Whimbrel","Eurasian Curlew","Bar-tailed Godwit","Dunlin","Avocet","Northern Lapwing")

Gulls<-c("Black-headed Gull","Black-legged Kittiwake","Lesser Black-backed Gull","Little Gull","Small gull species","tern/small gull","unidentified gull","unidentified large gull","unidentified small gull","Common Gull" ,"Great / Lesser Black-backed Gull","Herring Gull","Great Black-backed Gull","fulmar/gull","Fulmar / gull species",
         #ADDED 2020
         "fulmar/gull","Little Gull","Great / Lesser Black-backed Gull","tern/small gull")

Terns<-c("Sandwich Tern","Common/Arctic Tern","unidentified tern","Tern / small gull","Common Tern","Little Tern")

OtherBirds<-c("Common Kestrel","Eurasian Sparrowhawk","Grey Heron","unidentified songbird","unidentified bird","Eurasian Marsh Harrier","Common Wood Pigeon","Common Coot","Feral Pigeon",
              #added 2021-11-02
              "Great Egret")

Mammals <- c("Grey Seal", "Harbour Porpoise" ,"Harbour Seal","Minke Whale","unidentified marine mammal","unidentified pinniped (Grey/Harbour Seal)","White-beaked Dolphin","White-sided Dolphin","Bottlenose Dolphin", "Common Dolphin","Risso's Dolphin","Killer Whale")

assign('Divers',Divers,envir=globalenv())
assign('Cormorant_Gannets',Cormorant_Gannets,envir=globalenv())
assign('Anatidae',Anatidae,envir=globalenv())
assign('Geese_Shelducks',Geese_Shelducks,envir=globalenv())
assign('Skuas',Skuas,envir=globalenv())
assign('Auks',Auks,envir=globalenv())
assign('Grebes',Grebes,envir=globalenv())
assign('Waders',Waders,envir=globalenv())
assign('Gulls',Gulls,envir=globalenv())
assign('Terns',Terns,envir=globalenv())
assign('OtherBirds',OtherBirds,envir=globalenv())
assign('Mammals',Mammals,envir=globalenv())
}
ImportTaxonomicGroups(Species)

Code_Euring$TaxGroups<-ifelse(Code_Euring$English_name %in% Divers, "Divers",
                        ifelse(Code_Euring$English_name %in% Cormorant_Gannets, "Cormorant_Gannets",
                        ifelse(Code_Euring$English_name %in% Anatidae, "Anatidae",
                        ifelse(Code_Euring$English_name %in% Geese_Shelducks, "Geese_Shelducks",
                        ifelse(Code_Euring$English_name %in% Skuas, "Skuas",
                        ifelse(Code_Euring$English_name %in% Auks, "Auks",
                        ifelse(Code_Euring$English_name %in% Grebes, "Grebes",
                        ifelse(Code_Euring$English_name %in% Waders, "Waders",
                        ifelse(Code_Euring$English_name %in% Gulls, "Gulls",
                        ifelse(Code_Euring$English_name %in% Terns, "Terns",
                        ifelse(Code_Euring$English_name %in% OtherBirds, "OtherBirds", 
                        ifelse(Code_Euring$English_name %in% Mammals, "Mammals","Not included yet"))))))))))))
```


```{r, eval=FALSE, echo=FALSE}
Euring2<-seamonas::classify_taxonomic(my_data=Code_Euring,col_name = "English_name")
```

```{r, echo=FALSE, eval=FALSE}
Code_Euring$TaxGroups[1:5]<-'Divers'
Code_Euring$TaxGroups[6:14]<-'Grebes'
Code_Euring$TaxGroups[15:33]<-'Other_seabirds'
Code_Euring$TaxGroups[34:40]<-'Gannets_Cormorants'
Code_Euring$TaxGroups[41:57]<-'Other_species'
Code_Euring$TaxGroups[58:83]<-'Swan_Geese'
Code_Euring$TaxGroups[84:126]<-'Ducks_Shelducks'
Code_Euring$TaxGroups[127:176]<-'Other_species'
Code_Euring$TaxGroups[177:236]<-'Waders'
Code_Euring$TaxGroups[237:243]<-'Skuas'
Code_Euring$TaxGroups[244:273]<-'Gulls'
Code_Euring$TaxGroups[274:288]<-'Terns'
Code_Euring$TaxGroups[289:296]<-'Auks'
Code_Euring$TaxGroups[297:528]<-'Other_species'
Code_Euring$TaxGroups[529:546]<-'Marine_mammals'
Code_Euring$TaxGroups[547:549]<-'Other_species'
```

```{r}
Euring_code<-seamonas::Code_Euring
head(Euring_code)
```


## survey data

Data from a survey including longitude, latitude, timestamp

```{r, eval=FALSE}
survey_4326<-survey_4326
```

```{r}
head(survey_4326)
```

```{r}
ggplot2::ggplot()+
    ggplot2::geom_point(data=survey_4326,
                        ggplot2::aes(x=longitude, y= latitude, color=date),size = 1, shape = 16) 
```


## density survey

A data frame with survey data including observations, includes 2000 observations of 10 variables

```{r}
density_survey<-density_survey
```

```{r, eval=FALSE}
head(density_survey)
```

## density df


```{r,echo=FALSE, eval=FALSE}
## Create data frame
density_df<-data.frame(
             seq=(1:50),
             survey_id=rep(c(1, 2, 3, 4, 5), each = 10),
             longitude=c(seq(6, 8, length = 10),
                         seq(6, 8, length = 10),
                         seq(6, 8, length = 10),
                         seq(6, 8, length = 10),
                         seq(6, 8, length = 10)),
             latitude=c(seq(55.1,54, length = 10),
                        seq(55.1+0.1,54+0.1, length = 10),
                        seq(55.1+0.2,54+0.2, length = 10),
                        seq(55.1-0.2,54-0.2, length = 10),
                        seq(55.1-0.1,54-0.1, length = 10)),
             timestamps=c(
  head(seq.POSIXt(as.POSIXct("2018-04-01 09:00:00"), as.POSIXct("2018-04-01 18:00:00"), by="min"), 10),
  head(seq.POSIXt(as.POSIXct("2018-05-01 09:00:00"), as.POSIXct("2018-05-01 18:00:00"), by="min"), 10),
  head(seq.POSIXt(as.POSIXct("2018-06-01 09:00:00"), as.POSIXct("2018-06-01 18:00:00"), by="min"), 10),
  head(seq.POSIXt(as.POSIXct("2018-07-01 09:00:00"), as.POSIXct("2018-07-01 18:00:00"), by="min"), 10),
  head(seq.POSIXt(as.POSIXct("2018-08-01 09:00:00"), as.POSIXct("2018-08-01 18:00:00"), by="min"), 10))
)


density_df$year<-stringr::str_sub(density_df$timestamps, 1, 4)
density_df$month<-stringr::str_sub(density_df$timestamps, 6, 7)
density_df$day<-stringr::str_sub(density_df$timestamps, 9, 10)
density_df$date<-paste0(density_df$year,"-",density_df$month,"-",density_df$day)

#values 0 or 1
random1<-sample(0:1, 50, replace=TRUE)

#values 0 to 5.1

random2<-runif(50, min=0, max=5.1)

#make data frame

random3<-cbind(random1,random2)
random3<-as.data.frame(random3)


#substitute values
random4<-random3 %>%
  mutate(random5 = case_when(
    random1 == 0 ~ 0,
    random1 != 0 ~ random2,
    TRUE ~ random2))
#use as densities

density_df$densities<-as.numeric(random4$random5)

## Export
#save(density_df, file = "density_df.RData", version = 2)
```

A data frame with survey data includes 50 observations of 10 variables

```{r}
density_df<-density_df
```

```{r}
head(density_df)
```


## grid 5x5

A grid 5x5 in CRS 3035

```{r}
grid5x5_3035<-grid5x5_3035
```

```{r,eval=FALSE, eval=FALSE}
plot(grid5x5_3035)
```

```{r}
head(grid5x5_3035)
```

## surveyed grid

The surveyed grid, this means the grid cells were data was collected

```{r, eval=FALSE}
grid_surveyed<-grid_surveyed
```

```{r}
head(grid_surveyed)
```

# Survey grid

Spatial functions to transform the data

## transform_survey

A function to transform data collected in latitude and longitude from degrees, transform from 4326 to 3035

```{r}
survey_3035<-transform_survey(survey_data=survey_4326,
                 column_latitude='latitude',
                 column_longitude='longitude',
                 from_CRS=4326,
                 to_CRS=3035)
```


## subset_grid

Add values to geometries of the grid, each square is a geometry or polygon
There will be warning that variables are assumed to be spatially constant throughout all geometries
Will print a plot and return the grids with data

```{r, eval=FALSE}
grid_surveyed<-subset_grid(survey_grid=grid5x5_3035,
                      survey_data=survey_3035,
                      grid_identifier='grid_id')
```


plot to check grid

```{r, fig.height=4}
ggplot2::ggplot()+
  ggplot2::geom_sf(data = grid_surveyed, colour = "#42a921", fill= '#bde0fe',alpha=0.9)+
  ggplot2::geom_sf(data = survey_3035)+
  NULL
```




# Densities grid

A series of functions to go from a data frame of surveyed data to a grid

## transform_survey

A function to transform the CRS

```{r}
densities_3035<-transform_survey(survey_data=density_survey, 
                 column_latitude='latitude',
                 column_longitude='longitude',
                 from_CRS=4326,
                 to_CRS=3035)
```

## grid_densities

A function to keep only grids with data

```{r,eval=FALSE}
density_grid<-subset_density(density_survey=densities_3035,
                             column_density='densities',
                             survey_grid=grid_surveyed,
                             grid_identifier='grid_id')
```


Check the data

```{r, fig.height=4}
ggplot2::ggplot()+
  ggplot2::geom_sf(data = density_grid,mapping = ggplot2::aes(fill = mean_density), lwd = 0, colour = NA)
```


# Plots

Functions to facilitate plotting on the same style

```{r}
library(sf)
library(GermanNorthSea)
library(ggplot2)
```

## Points

Base map

Define parameters, for this example the CRS **4326** is to be used. 

```{r}
my_CRS<-4326
Europa<-sf::st_transform(German_land, my_CRS)
EEZ<-sf::st_transform(German_EEZ, my_CRS)
color_land='#f7bf54'
color_water='#3668b4'
xval<-c(3,9)
yval<-c(53,56)
```

```{r}
Base_map<-ggplot2::ggplot()+
    # maps
    ggplot2::geom_sf(data = EEZ, colour = 'black', fill = color_water)+
    ggplot2::geom_sf(data = Europa, colour = 'black', fill = color_land)+ 
    ggplot2::coord_sf(xlim = xval, ylim = yval)+

    NULL
Base_map
```



### Add dots

Data from **density_df** will be used, define shape, size and labels.

```{r}
class0<-0
class1<-1
class2<-2.5
class3<-5
```

```{r}
density_df<-density_df %>% 
    dplyr::mutate(density_class = dplyr::case_when(
      densities ==  class0 ~ as.character("group0"),
      densities >=  class0 & densities <=  class1 ~ as.character("group1"),
      densities >=  class1 & densities <=  class2 ~ as.character("group2"),
      densities >=  class2 & densities <=  class3 ~ as.character("group3"),
      densities >=  class3 ~ as.character("group4"),
      TRUE~"U"))
```

```{r}
density_sizes<-c("group0"=0.5, "group1"=1,"group2"=2, "group3"=3, "group4"=5)
density_shapes<-c("group0"=3, "group1"=21,"group2"=21, "group3"=21, "group4"=21)
density_labels<-c('0','> 0-1','> 1-2.5','> 2.5-5','> 5')
```

```{r}
density_wmap<-Base_map+
  geom_point(data=density_df,
          aes(x = longitude,
              y= latitude,
              shape = density_class,
              size= density_class),
              fill= "#d00000")+
  
  scale_shape_manual(values = density_shapes,labels=density_labels)+
  scale_size_manual(values =  density_sizes,labels=density_labels)
density_wmap
```

```{r, eval=FALSE, echo=FALSE}
#Export
#save(density_wmap, file = "density_wmap.RData", version = 2)
```

### add_legend

A function to add the legend inside the plot
```{r, eval=FALSE,echo=FALSE}
add_legend<-function(plot_wbreaks=plot_wbreaks,
                     legxy=legxy){

  plot_wlegend<-plot_wbreaks+
    ggplot2::theme(
      legend.position = legxy,
      legend.title = ggplot2::element_blank(),
      legend.text= ggplot2::element_text(size=10,color="#343a40",family='sans'),
      legend.spacing.y =  ggplot2::unit(0.01, 'cm'),
      legend.spacing.x =  ggplot2::unit(0.2, 'cm'),

      legend.background = ggplot2::element_rect(fill='transparent',colour ="transparent"),
      legend.box.background = ggplot2::element_rect(fill='transparent',colour ="transparent"),
      legend.key = ggplot2::element_rect(fill = "transparent", colour = "transparent"),

      #size
      legend.key.size =  ggplot2::unit(0.7, 'cm')
    )

  return(plot_wlegend)
}
```

```{r}
density_wlegend<-add_legend(
  plot_wbreaks=density_wmap,
  legxy=c(0.11, 0.21))
```

```{r, message=FALSE, warning=FALSE}
density_wlegend<-density_wlegend+
  theme(legend.key.size = ggplot2::unit(0.4, "cm"))+
  ggplot2::annotate(geom="text",
                      x=3.0, y=54.0,
                      label=expression(atop("Density"), paste("[ind/k", m^2,"]")),
                      color="#343a40",hjust = 0)
density_wlegend
```

### add_theme

```{r, eval=FALSE, echo=FALSE}
add_theme<-function(plot_wlegend=plot_wlegend){

  plot_wtheme<-plot_wlegend+
    ggplot2::xlab('Longitude')+
    ggplot2::ylab('Latitude')+
    
    # theme
    ggplot2::theme(

      #panel
      panel.border = ggplot2::element_rect(colour = "black", fill=NA, size=1),
      panel.grid.major = ggplot2::element_blank(),
      panel.grid.minor = ggplot2::element_blank(),
      panel.background = ggplot2::element_blank())

  return(plot_wtheme)
}
```

```{r, message=FALSE, warning=FALSE}
density_wtheme<-add_theme(plot_wlegend = density_wlegend)
density_wtheme
```


# Grid

Base data

```{r}
my_CRS<-3035
Europa<-sf::st_transform(GermanNorthSea::German_land, my_CRS)
EEZ<-sf::st_transform(GermanNorthSea::German_EEZ, my_CRS)
color_land='#f7bf54'
color_water='#3668b4'
xval<-c(3910000,4250000)
yval<-c(3380000,3680000)
```

Provide density grid and mean densities

```{r}
density_plot<-ggplot2::ggplot()+
    # maps
    ggplot2::geom_sf(data = EEZ, colour = 'black', fill = color_water)+
    ggplot2::geom_sf(data = Europa, colour = 'black', fill = color_land)+
    ggplot2::geom_sf(data = density_grid,
                     mapping = ggplot2::aes(fill = mean_density), lwd = 0, colour = NA) + 
  ggplot2::coord_sf(xlim = xval, ylim = yval)+

    NULL
density_plot
```


## add_breaks

A function to add color palette and breaks in the legend, will vary with the species density estimates

```{r}
add_breaks<-function(density_plot=density_plot,
                     legendbreaks=legendbreaks,
                     legendlimits=legendlimits,
                     legendlabels=legendlabels){

  plot_wbreaks<-density_plot+
    ggplot2::scale_fill_gradientn(

      colours = c("#f1faee", #0 white
                  "#4ea8de", #0.10 lightblue
                  '#1e6091', #0.20 darkblue
                  "#34a0a4", #0.30 darkgreen
                  '#76c893', #0.40 lightgreen
                  '#d9ed92', #0.50 darkyellow
                  "#ffea00", #0.60 bringyellow
                  "#f1ba0a", #0.70 organge
                  '#ff4800', #0.80 darkorgance
                  "#e01e37", #0.90 lightred
                  '#6e1423'  #1.0 darkred
      ),
      #the values are according to the size of the scale (100%), not values of the column
      values = c(0, 0.10,0.20,0.30,0.40,0.50,0.60,0.70,0.80,0.90,1.0),

      limits=legendlimits,
      breaks=legendbreaks,
      labels=legendlabels)+
    ggplot2::guides(fill = ggplot2::guide_colourbar(ticks = FALSE))
  return(plot_wbreaks)
}
```

```{r, eval=TRUE}
plot_wbreaks<-add_breaks(density_plot=density_plot,
                     legendbreaks=c(0,0.5,1,1.5,2.0),
                     legendlimits=c(0,2),
                     legendlabels=c('     0','   0.5','   1.0','   1.5','> 2.0'))
plot_wbreaks
```



## add_legend

A function to add the legend inside the plot

```{r}
plot_wlegend<-add_legend(
  plot_wbreaks=plot_wbreaks,
  legxy=c(0.11, 0.21))
```

```{r, message=FALSE, warning=FALSE}
plot_wlegend<-plot_wlegend+
  ggplot2::theme(legend.key.size =  ggplot2::unit(0.7, 'cm'))+
  ggplot2::annotate(geom="text",
                      x=3905000, y=3510000,
                      label=expression(atop("Density"), paste("[ind/k", m^2,"]")),
                      color="#343a40",hjust = 0)
plot_wlegend
```


##  add_theme

A function to define the theme

```{r, eval=FALSE, echo=FALSE}
add_theme<-function(plot_wlegend=plot_wlegend){

  plot_wtheme<-plot_wlegend+
    ggplot2::xlab('Longitude')+
    ggplot2::ylab('Latitude')+
    ggplot2::scale_x_continuous(labels = function(x) paste0(x, '\u00B0', "W")) +
    ggplot2::scale_y_continuous(labels = function(x) paste0(x, '\u00B0', "N"))+

    # theme
    ggplot2::theme(

      #panel
      panel.border = ggplot2::element_rect(colour = "black", fill=NA, size=1),
      panel.grid.major = ggplot2::element_blank(),
      panel.grid.minor = ggplot2::element_blank(),
      panel.background = ggplot2::element_blank())

  return(plot_wtheme)
}
```

```{r, message=FALSE, warning=FALSE}
plot_wtheme<-add_theme(plot_wlegend = plot_wlegend)
plot_wtheme
```







